Bricking up backdoors in Intrusion Detection Systems

### 2019-09-16 paper/paper.tex
Add missing minus to figure

For reproducibility, we make all our code publicly available at \url{https://github.com/CN-TU/ids-backdoor}.

@@

There have been several works with the aim of increasing robustness of decision trees against attacks. \cite{biggio_bagging_2011} proposed a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset. Their method is applicable to \glspl{dt}. However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work. \cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks. \cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
%\todo{Max: Unnecessary or should be moved to Related Work?}
Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
We developed a defense approach specifically for \gls{rf} classifiers.
The concept behind our pruning defense is that leaves that are never used by samples in the validation set might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation set should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
\item Like (2) but additionally use depth to decide when
%two % AH could be more than two
leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
\end{enumerate}

@@

For \cic{} we get similar results, but pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains). We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$. Thus, backdoored flows cannot always be sharply distinguished from regular flows. We find that pruning only harmless leaves is very beneficial for \unsw{} but not for \cic{}. However, considering depth when making the decision which leaf to prune first, leads to the backdoor being removed considerably earlier in the pruning process.

>>>>

For reproducibility, we make our code, data and figures publicly available at \url{https://github.com/CN-TU/ids-backdoor}.

@@

There have been several works with the aim of increasing robustness of decision trees against attacks. \cite{biggio_bagging_2011} proposed a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset. Their method is applicable to \glspl{dt}. However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work. \cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks. \cite{russu_secure_2016} makes SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
%\todo{Max: Unnecessary or should be moved to Related Work?}
Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
We developed a defense approach specifically for \gls{rf} classifiers.
The concept behind our pruning defense is that leaves that are never used by samples in the validation set might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation set should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``benign'' leaves: We assume that attackers want malicious samples appear benign.
\item Like (2) but additionally use depth to decide when
%two % AH could be more than two
leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
\end{enumerate}

@@

For \cic{} we get similar results, but pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains). We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$. Thus, backdoored flows cannot always be sharply distinguished from regular flows. We find that pruning only benign leaves is very beneficial for \unsw{} but not for \cic{}. However, considering depth when making the decision which leaf to prune first, leads to the backdoor being removed considerably earlier in the pruning process.
